Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-02-17 docs/papers/roofline_model/gocean_roofline.tex
Roofline: add perf-plot of Continuity kernel for various SIMD optimisations

Peak absolute performance is obtained for the 256 problem size with
SSE instructions. The 512 problem size spills from L3 cache and
becomes bound by the bandwidth to main memory: SIMD vectorisation
gives no performance benefit. The fact that best performance is
obtained when the problem size fits within L3 indicates that there are
other overheads that dominate when the problem size is small enough to
fit within L2. This is possibly because the inner, SIMD-vectorised
loop then only has a trip-count of 35. (This is not 32 because of the way
the simulation domain is set-up; it is the {\em internal} region of
the domain that has dimension $32^2$.) {\bf Remainder and peel loops.}

@@

>>>>

Peak absolute performance is obtained for the $256^2$ problem size
with SSE instructions. The 512 problem size spills from L3 cache and
becomes bound by the bandwidth to main memory: SIMD vectorisation
gives no performance benefit. The fact that best performance is
obtained when the problem size fits within L3 indicates that there are
other overheads that dominate when the problem size is small enough to
fit within L2. We checked the efficiency of the SIMD vectorised code
by looking at the compiler diagnostic output. On the basis of this we
added Intel-specific directives to the code in order to specify that
the arrays were aligned on 64-byte boundaries. We also manually coded
the 'peel loop' (iterations required to reach a 64-byte
boundary). Having done this the compiler reported that no peel or
remainder loops were generated and the resulting (SSE) vectorised loop
had an estimated speed-up of 1.9. (This estimated speed-up was
previously 1.6; in agreement with the speed-up we measured.) After
making these changes the performance of the Continuity kernel with the
$32^2$ domain increased from 4.22 to 4.57 GFLOPs.

@@

\begin{figure}
  \centering
  \includegraphics[width=120mm]{continuity_opt}
  \caption{Performance of the Continuity kernel with optimisations
    intended to aid SIMD vectorisation. Results are for v.16 of the
    Intel compiler on a Xeon E5-1620 v.2 CPU.}
  \label{FIG_cont_vec}
\end{figure}
